```{r}
library(plyr)
library(ggplot2)
library(chron)
library(knitr)
```
#Loading and preprocessing the data
```{r}
setwd("H:/Xiao Wang/Coursera/Course 5/week2")
activity<-read.csv("activity.csv", header = TRUE)
head(activity)
```
#Q1 What is mean total number of steps taken per day?
#For this part of the assignment, you can ignore the missing values in the dataset.
#Make a histogram of the total number of steps taken each day
#Calculate and report the mean and median total number of steps taken per day

#Make a histogram
```{r}
activity1<-activity[(!is.na(activity$steps)),]
tstep<-with(activity1,tapply(steps,date,sum,na.rm=TRUE))
tstepf<-data.frame(date=names(tstep),sum=tstep)
tstepf<-tstepf[(!is.na(tstepf$sum)),]
hist(tstepf$sum,xlab="number of steps",main = "Histogram of the total number of steps taken each day")
```
#mean and median of steps
```{r}
as.integer(mean(tstepf$sum))
as.integer(median(tstepf$sum))
```

#Q2 What is the average daily activity pattern?
#a.Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) 
#and the average number of steps taken, averaged across all days (y-axis)
#b.Which 5-minute interval, on average across all the days in the dataset, 
#contains the maximum number of steps?

#a
```{r}
tstepi<-with(activity1,tapply(steps,interval,mean,na.rm=TRUE))
plot(as.numeric(names(tstepi)),tstepi,xlab="interval",ylab="mean steps",type="l")
```

#b
```{r}
tstepi<-data.frame(interval=names(tstepi),mean=tstepi)
tstepi$mean<-as.numeric(tstepi$mean)
summary(tstepi)
subset(tstepi,mean>206)
```

#Q3 Imputing missing values
#Note that there are a number of days/intervals where there are missing values 
#(coded as NA). The presence of missing days may introduce bias into some 
#calculations or summaries of the data.

#a. Calculate and report the total number of missing values in the dataset 
#   (i.e. the total number of rows with NAs)
#b. Devise a strategy for filling in all of the missing values in the dataset. 
#   The strategy does not need to be sophisticated. For example, you could use the 
#   mean/median for that day, or the mean for that 5-minute interval, etc.
#c. Create a new dataset that is equal to the original dataset but 
#   with the missing data filled in.
#d. Make a histogram of the total number of steps taken each day, and Calculate 
#   and report the mean and median total number of steps taken per day. 
#   Do these values differ from the estimates from the first part of 
#   the assignment? What is the impact of imputing missing data on the 
#   estimates of the total daily number of steps?

#a
```{r}
sum(is.na(activity$steps))
```

#b
```{r}
activity$steps<-as.character(activity$steps)
na<-subset(activity,is.na(activity$steps)=="TRUE",select=c(date,interval))
intmean<-merge(tstepi,na)
```

#c
```{r}
names(intmean)[2]<-"steps"
no_na<-subset(activity,is.na(activity$steps)=="FALSE")
```
#reorder the column
```{r}
no_na<-no_na[,c(2,3,1)]
intmean<-intmean[,c(3,1,2)]
rbd<-rbind(no_na,intmean)
rbd$date<-as.Date(rbd$date,tryFormats = c("%Y-%m-%d", "%Y/%m/%d"))
rbd$steps<-as.numeric(rbd$steps)
rbd$interval<-as.numeric(rbd$interval)
```
#d
```{r}
tstepd<-with(rbd,tapply(steps,date,sum,na.rm=TRUE))
tstepd<-data.frame(date=names(tstepd),sum=tstepd)
hist(tstepd$sum,xlab="number of steps",main = "Histogram of the total number of steps taken each day (Imputed Data)")

mean(tstepd$sum)
median(tstepd$sum)
```
#The mean remains the same as prior to imputation, while the median value increased slightly.

#Q4 
# Are there differences in activity patterns between weekdays and weekends?
#   For this part the weekdays() function may be of some help here. Use the dataset with 
#   the filled-in missing values for this part.
# Create a new factor variable in the dataset with two levels 
#   -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
# Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute 
#   interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r}
rbd$weekday<-ifelse(weekdays(rbd$date)==c("Saturday","Sunday"),"weekend","weekday")
new<-arrange(rbd,interval)
weekday<-subset(new,weekday=="weekday")
weekend<-subset(new,weekday=="weekend")
weekdayplt<-with(weekday,tapply(steps,interval,mean,na.rm=TRUE))
weekendplt<-with(weekend,tapply(steps,interval,mean,na.rm=TRUE))

par(mfrow=c(2,1))
plot(as.numeric(names(weekdayplt)),weekdayplt,xlab="interval",ylab="mean steps", main = "Weekdays Activity Pattern", type="l")
plot(as.numeric(names(weekendplt)),weekendplt,xlab="interval",ylab="mean steps", main = "Weekends Activity Pattern", type="l")

```